Language Skills in an Ethnically Segmented Labour market: Estonia 1989 -- 2012

### 2018-07-20 response-to-reviewers.tex

\begin{response}
  We have corrected the different references.
\end{response}

\item Page 2, line 58: Not the unemployment rate, but the probability of
  being unemployed is reduced.

  \begin{response}
    We now use ``unemployment probability'' when talking about
    individual outcomes and leave ``unemployment rate'' solely for
    describing group outcomes.
  \end{response}

\item Page 3, lines 46-50: You are using a comma as the decimal separator, but in all your tables you are using a point. Please use a point everywhere.

  \begin{response}
    We now consistently use period as decimal separator and comma as
    thousands' separator.
  \end{response}

\end{enumerate}

@@

\section{Referee: 2}

\subsection{Additional Questions:}

\subsubsection*{4. Results}

\begin{enumerate}
\item For this papers findings section in the abstract they imply that
  for all men fluency in Estonian is related to 5\% lower unemployment
  and no income premium.  From Table 2 and 3 it appears that this
  should be for only Russian males. English skills are associated with
  a large income premium and no unemployment. From Table 2 and 3 this
  appears to be more accurately just Russians, as the Estonians appear
  to have a significant negative correlation with unemployment.  This
  is supported by what is said later in the Discussion and Concluding
  Remarks section in what they state is their most intriguing
  observation.  Claiming this to be the most intriguing is interesting
  as it seems to support some what previous literature has stated and
  one of the extensions that this paper is making the inclusion of
  women and namely how their effects differ from males and previous
  literature.

  \begin{response}
    No response neede here.
  \end{response}

\item In the Discussion and Concluding Remarks section they state that
  “The lower end labour market, associated with frequent unemployment,
  is better integrated and those who are fluent in the language may
  expect both less unemployment and higher wage. However, men at
  better paid and more stable jobs gain little form Estonian” (the
  language I presume) This conclusion felt surprising when reading the
  paper as it did not feel fully fleshed out in the results section
  how this was observed. Tying this to the results would be helpful
  and make it clearer exactly where these results come from.

  \begin{response}
    We have completely re-phrased this paragraph and rely more on our
    results (and literature) now.
  \end{response}

\item In Table 3 I am perplexed how the number of observations increases as more restrictions are added to the model. How could running the same model but including additional workplace controls increase the observations from 7978 to 8233 for men and 9697 to 9895 for women? It makes me somewhat cautious of the results—though it could just be a typo. Could the models be run using the same sample for each column?
\debug{These typos got fixed.}
\end{enumerate}

### 2018-07-20 paper/ijm.tex
the fuzzy paragraph about integration at the low end of job market addressed

We document that all analyzed language skills are associated with
unemployment and wage, however in a different way. Our most
intriguing observation is that the largest group of
Russian men, high school graduates, earn virtually no income premium if fluent in Estonian,
the majority and the official language of the country, while ability
to speak the language is related to a substantially
lower unemployment probability.
Our analysis supports \citet{YaoandOurs2015}, and \citet{Lindemann2013} findings that
the labour market is segmented between ethnic and gender lines.
However, we find that ethnic segregation is less prevalent in the male
labour market.
This suggests for majority of Russian speaking men, Estonian
language gives a distinct advantage by facilitating hiring by
Estonian-dominated employers.  However, our outcome also supports 
\citet{leppik+vihalemm2015JofBaltStud} conclusion that
these
jobs are associated with little career mobility.  There is also
anecdotal evidence suggesting that offices are
typically manned by ethnic Estonians while manufacturing units have a
large share of Russian workers.
One can speculate that this outcome is related to a substantial population share
with negative attitudes toward Russians \citep{korts2009JofBaltStud}.
On a more positive tone, we find
that this may not be the case for highly educated workers.

The ethnic segregation seems to be less of an issue for women.
Unfortunately, neither \citet{Toomet2011} nor
\citet{leppik+vihalemm2015JofBaltStud} analyse men and women
separately.  We find some evidence that female labor market is less
segregated along ethnic lines, and women are more frequently employed in jobs in which they need Estonian
language more, such as working in a direct contact with customers.

We also find evidence that over time Estonian is becoming more
important in the female labour market.  Unfortunately, the current data
does not let us to assess whether is due to decreasing segregation, or
economy--wide rise of the importance of Estonian language.  No such trend is visible for
men suggesting that 20 years--almost a generation after the collapse of Soviet Union, the
labor market integration is still not complete.

The other results are more in line with other studies.
English language is related to a substantial income premium but
virtually to no effect on unemployment.  This is similar to findings by
\citet{ginsburgh+prieto-rodriguez2011ILRR} and
\citet{beblavy+2016CEPS} who show that English is more important at better paid
jobs.  In contrast, our outcomes suggest that English has little use
at unstable jobs, at these jobs where workers frequently move between
employment and unemployment.
Note also that for
English, we are able to explain a substantial part of the raw effect
by individual characteristics, suggesting that the true causal effect
may be smaller than our estimates.

The study has a number of limitations.  Although we believe that our
estimates are good indicators of a causal relationship, we cannot 
claim this as our data do not contain strong enough instruments.
In particular, as individual characteristics are able to explain a
large part of English returns, we suppose part of the remaining effect
may also related to an ability bias.  Another weak point in the current
data is using the self-reported language
skills, although this is the standard practice in the literature.  As language test data becomes available, we may be able to
decrease the measurement error in the future.

>>>>

We document that all analyzed language skills are associated with
unemployment and wage, however in a different way. Our most
intriguing observation is that the largest group of
Russian men, high school graduates, earn virtually no income premium
from being fluent in Estonian,
the majority and the official language of the country, while ability
to speak the language is related to a substantially
lower unemployment probability.
Our analysis supports \citet{YaoandOurs2015}, and \citet{Lindemann2013} findings that
the labour market is segmented between ethnic and gender lines.
However, we find that ethnic segregation is more prevalent in the male
labor market.
It is true that for majority of Russian speaking men, Estonian
language gives a distinct advantage by facilitating hiring by
Estonian-dominated employers.  However, our outcome also supports 
\citet{leppik+vihalemm2015JofBaltStud} conclusion that
these
jobs are associated with little career mobility.  There is also
anecdotal evidence suggesting that offices are
typically manned by ethnic Estonians while manufacturing units have a
large share of Russian workers.
One can speculate that this outcome is related to a substantial population share
with negative attitudes toward Russians \citep{korts2009JofBaltStud}.
On a more positive tone, we find
that this may not be the case for highly educated workers.

The ethnic segregation seems to be less of an issue for women.
Unfortunately, neither \citet{Toomet2011} nor
\citet{leppik+vihalemm2015JofBaltStud} analyse men and women
separately.  We find some evidence that female labor market is less
segregated along ethnic lines, women are also more frequently employed
in jobs where they need Estonian
language more, possibly because of being in a direct contact with customers.

We also find evidence that over time Estonian is becoming more
important in the female labour market.  Unfortunately, the current data
does not let us assess whether this is due to decreasing segregation,
or due to
economy--wide rise of the importance of Estonian language.  No such trend is visible for
men suggesting that 20 years---almost a generation---after the collapse of Soviet Union, the
labor market integration is still not complete.

The other results are more in line with other studies.
English language is related to a substantial income premium but
to virtually no effect on unemployment.  This is similar to findings by
\citet{ginsburgh+prieto-rodriguez2011ILRR} and
\citet{fabo+2017E} who show that English is more important at better paid
jobs.  In contrast, our outcomes suggest that English has little use
at unstable jobs, at these jobs where workers frequently move between
employment and unemployment.
Note also that for
English we are able to explain a substantial part of the raw effect
by individual characteristics, suggesting that the true causal effect
may be smaller than our estimates.

The study has a number of limitations.  Although we believe that our
estimates are good indicators of a causal relationship, it is still
somewhat uncertain as our data do not contain strong instruments.
In particular, as individual characteristics are able to explain a
large part of English returns, we suppose part of the remaining effect
may also related to the ability bias.  Another weak point in the current
data is using the self-reported language
skills, although this is the standard practice in the literature.  As language test data becomes available, we may be able to
decrease the measurement error in the future.
